---
title: "p8105_hw3_sw3879"
author: "Siqing Wang"
date: "2023-10-12"
output: github_document
---

Setting up libraries
```{r setup, message = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## Problem 1

### Loadiing the dataset for Q1
```{r}
library(p8105.datasets)
data("instacart")

instacart = instacart |> as.tibble() |> janitor::clean_names()
```

### Describing the dataset 
The `instacart` dataset as `r nrow(instacart)` rows and `r ncol(instacart)` columns. Important variables include unique identifier of customer, order date, days since previous order, and specific information about the product purchased. There are a total of `r instacart |> select(order_id) |> n_distinct()` orders recorded, and `r instacart |> select(product_id) |> n_distinct()` recorded. These products range from `r instacart |> select(department_id) |> n_distinct()` different departments. 

### Answering questions 
```{r}
instacart |> count(aisle) |> arrange(desc(n))
```

There are `r instacart |> select(aisle_id) |> n_distinct()` aisles, and the most ordered items are from fresh vegetables and fresh fruits. 



